### 2017-03-10 thesis/results/results.tex
Changed all the data set names and removed "'s" - tmp commit

The obesity associated genetic signatures are central to this project in order to clarify the relationship between \gls{bmi} and cancer.
Firstly in this chapter, the previously identified obesity associated signatures from the studies conducted by \citet{Creighton2012} and \citet{Fuentes-Mattei2014} were examined in turn to judge the agreement of these signatures with the sample \gls{bmi}/\gls{bmi} status, as presented in thier results.
Secondly, novel obesity associated genetic signatures were identified in the \citet{Creighton2012} data set and was compared with Creighton's and Fuentes-Mattei's  signatures.
Lastly, the presence of common genes or pathways that were associated with obesity in multiple types of cancer was explored using the \gls{icgc} cancer data sets.

@@

Obesity metagene was created using the obesity associated genetic signatures from \citet{Creighton2012} study.
% As the genetic signatures were described as gene probe IDs, the gene probes had to be converted into gene symbols (see \cref{sec:data}).
% This resulted in 644 unique gene symbols, compared to 799 gene probes IDs in the original signature.
There was no description about the normalisation method used by Creighton's group when they first analysed their data, so the more reliable \gls{rma} method was used to  normalise the Creighton data set.
The obesity metagene was plotted in a heatmap with the sample gene expression  to check whether the metagene scores were in accordance with the overall gene expression of the samples (\cref{fig:crmetaheat}).

\begin{figure}[htb]
	\centering
	\includegraphics[page=3,width=0.8\linewidth]{results1/creighton_mg_heatmap1}
	\caption[Creighton's metagene and sample gene expression in Creighton's data]{Heatmap showing Creighton's obesity associated metagene with the gene expressions of obesity associated genes in Creighton's data.
	Level of expression is represented in the top right histogram, where low and high gene expression were colour-coded with blue and red, respectively.
	Each row of the heatmap represents a gene from the obesity associated genetic signature, and each column of the heatmap represents a sample from Creighton's data.
	The obesity associated metagene scores of the samples are shown in a separate row at top of the heatmap, and the tree diagram of the heirarchical clustering of the genes is shown to the left of the heatmap.
	For clarity, the metagene scores were flipped in order to match the results from \citet{Creighton2012} paper.}
	\label{fig:crmetaheat}
\end{figure}

@@

\begin{figure}[htp!]
	\centering
	\includegraphics[page=4,width=0.45\linewidth]{results1/creighton_mg_heatmap1}
	\hfill
	\includegraphics[page=5,width=0.45\linewidth]{results1/creighton_mg_heatmap1}
	\caption[Creighton's metagene and sample \gls{bmi}/\gls{bmi} status in Creighton's data]{Box plot and scatter plot showing the association of Creighton's obesity associated metagene with sample \gls{bmi} status and \gls{bmi} from Creighton's data, respectively.
	In the box plot, the p-values above the groups represent the statistical significance of the association of the metagene with the overweight or obese group compared with the normal weight group.
	The \gls{anova} p-value shows the statistical significance of the association of the metagene with the sample \gls{bmi} groups.
	In the scatter plot, $R^2$- and p-values describe the adjusted coefficient of determination of the regression line and the statistical significance of the linear model used to draw the regression line, respectively.}
	\label{fig:crmetaboxplot}
\end{figure}

\cref{fig:crmetaboxplot} clearly showed that the obesity associated metagene from \citet{Creighton2012} study significantly associated with the sample \gls{bmi} status, as well as sample \gls{bmi} value (include p-value and r-squared).
It should be noted here that the obesity associated metagene significantly associated with the samples that were obese, but not with the samples that were overweight.
This was due to the fact that the obesity associated genetic signatures were originally identified from the comparison of the samples that were obese with the samples that were not obese, and therefore the metagene scores were significant with the obese group, but not with the overweight group.
In addition to this, although the regression line in the scatter plot showed statistically significant association with sample \gls{bmi}, the sample \gls{bmi} values seemed to be randomly dispersed across the metagene scores.
This suggested that perhaps Creighton's obesity metagene was not associated with sample \gls{bmi} as strongly as expected.
\\

\noindent
Now that the association of the obesity metagene from \citet{Creighton2012} study was established in Creighton's data set, the obesity associated signature was generated in the \gls{icgc} cancer data.
The direction of the obesity associated metagene was checked in the Creighton's data first, so that high metagene scores reflected high sample \gls{bmi} and low metagene scores reflected low sample \gls{bmi} (\cref{fig:crmetaboxplot}).
The transformation matrix was then created in Creighton's data, as described in \cref{sub:svd}.

All of the \gls{icgc} data were normalised as described in \cref{ssub:rna_seq_data}.
Before the transformation matrix was applied to the log$_{10}$-normalised cancer data, the suitability of standardised data or untouched (non-standardised) data was determined (see \cref{sec:metagenes_created_from_raw_data_vs_standardised_data_icgc}).
From these results, the standardised data was found to be the most suitable to use for matrix transformation.
The transformation matrix was applied to each cancer data set in turn to obtain Creighton's obesity metagenes for each of the data set.
Each obesity metagenes were plotted in a heatmap with the corresponding data set in which the metagene was taken from (\cref{fig:crmetaicgc}; \cref{sec:rest_of_the_icgc_cancer_heatmap_resutls}).
These heatmaps confirmed that the obesity associated metagene derived from Creighton's data was able to capture the overall gene expression pattern, where the metagene scores reflected the expression levels of majority of the genes in the signature.
As before, association of the obesity metagene with the sample \gls{bmi} and \gls{bmi} status was checked in their respective cancer data set (\cref{fig:crmetaicgc}; \cref{sec:rest_of_the_icgc_cancer_heatmap_resutls}).
Out of all the cancer types, only \gls{blca} data set showed significant association of Creighton's obesity metagene with the overweight group (but not with the obese group).
However, neither the \gls{anova} p-value nor the regression line in the scatter plot were statistically significant, suggesting that this apparent association with the overweight group was not due to the Creighton's obesity metagene (?).

\begin{figure}[htp!]
	\centering
	\includegraphics[page=3,width=0.8\linewidth]{results1/crtcga_std}\\
	\vspace{1em}
	\includegraphics[page=4,width=0.45\linewidth]{results1/crtcga_std}
	\hfill
	\includegraphics[page=5,width=0.45\linewidth]{results1/crtcga_std}
	\caption[Creighton's metagene in \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of Creighton's obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from \gls{icgc} \gls{blca} data, respectively.
	The results for other cancer types are in \cref{app:a}.
	Scales, p-values and $R^2$-value are as described in previous figures.}
	\label{fig:crmetaicgc}
\end{figure}

There could be several reasons for the apparent lack of association of the metagene with sample \gls{bmi}/\gls{bmi} status.
First, transformation matrix was derived from Creighton's microarray data, but the \gls{icgc} cancer data were \gls{rnaseq} data.
Though the log$_{10}$ normalisation and standardisation of the data was the most appropriate adjustment to be made to the \gls{rnaseq} data, this adjustment was not equivalent to the \gls{rma} normalisation method that was used on the microarray data.
Secondly, none of the \gls{icgc} cancer originated from breast as in Creighton's data.
Since the obesity associated signature was identified in breast cancer data, the signature may be specific to breast cancer and may not be suitable in other cancer types.
\\

\noindent
To check whether the metagene was specific to breast cancer microarray data, the same transformation matrix was applied to breast cancer microarray data from \citet{Print2016} study (referred to as Print's data).
Print's data was normalised with \gls{rma} method and the transformation matrix was applied to the normalised data to obtain the metagene.
The metagene was again compared with the gene expression of the samples with a heatmap and the association of the metagene with the sample \gls{bmi}/\gls{bmi} status was examined with box and scatter plots (\cref{fig:crmetaprint}).

The obesity metagene managed to reflect the overall gene expression of the samples in Print's data.
However, as with \gls{icgc} cancer data, the obesity metagene scores did not significantly associate with the sample \gls{bmi} or \gls{bmi} status.
These results confirmed that the lack of association of the obesity metagene from Creighton's data was not due to the technology in which the data was gathered (mcroarray or \gls{rnaseq}), nor the cancer type in which the transformation matrix was applied to.

\begin{figure}[htp!]
	\centering
	\includegraphics[width=0.8\linewidth,page=3]{results1/cris_cr_trans_meta}\\
	\vspace{1em}
	\includegraphics[width=0.45\linewidth,page=4]{results1/cris_cr_trans_meta}
	\hfill
	\includegraphics[width=0.45\linewidth,page=5]{results1/cris_cr_trans_meta}
	\caption[Creighton's metagene in Print's breast cancer data]{Heatmap, box plot and scatter plot showing the association of Creighton's obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from Print's breast cancer data, respectively.
	Scales, p-values and $R^2$-value are as described in previous figures.}
	\label{fig:crmetaprint}
\end{figure}

All together, all of these results suggest that the obesity metagene identified in the study conducted by \citet{Creighton2012} associated with sample \gls{bmi} and \gls{bmi} status only in Creighton's data.
The lack of association with \gls{bmi} was not due to the type of technology platform in which the data was gathered, as neither the \gls{icgc} \gls{rnaseq} data nor Print's microarray data showed no significant association with the obesity metagene.
Furthermore, Creighton's metagene was not dependent on the cancer type since the obesity associated genetic signature did not show significant association in neither the \gls{icgc} cancer data nor in Print's breast cancer data.

One possible reason why Creighton's obesity metagene did not show significant association with other data sets could be because the genetic signature was in fact not an obesity specific signature, but a signature that was detected due to another clinical variable.
Another reason for this apparent lack of association could be that the genetic signature was too specific to the data and was not a broad obesity associated genetic signature, but an obesity associated signature specifically for Creighton's data set.

@@

Since Creighton's obesity associated genetic signature was not associated with sample \gls{bmi} or \gls{bmi} status in other data sets, Fuentes-Mattei's (FM) obesity associated genetic signature was examined to see whether FM's obesity metagene associated with sample \gls{bmi} or \gls{bmi} status.
Since FM's data set did not have sample \gls{bmi} information, FM's obesity metagene was not able to be compared with the sample \gls{bmi} or \gls{bmi} status in the original FM's data.
However, the transformation matrix was still applied to other cancer data to see whether FM's obesity metagene associated with sample \gls{bmi}/\gls{bmi} status

Firstly, FM's data was normalised with the \gls{rma} method and \gls{svd} was applied to the normalised data to get the transformation matrix.
The transformation matrix was used to transform the \gls{rma}-normalised Creighton's data to extract FM's obesity metagene scores in Creighton's data.
FM's obesity metagene scores were compared with gene expressions and sample \gls{bmi}/\gls{bmi} status in Creighton's data, as shown in \cref{fig:fmmetacr}.
Clearly, like with Creighton's obesity metagene, FM's obesity metagene was reflective of the overall gene expression of the samples, but did not associate with the sample \gls{bmi} or \gls{bmi} status.

\begin{figure}[htp!]
	\centering
	\includegraphics[page=3,width=0.8\linewidth]{results1/cr_fm_meta}\\
	\vspace{1em}
	\includegraphics[page=4,width=0.45\linewidth]{results1/cr_fm_meta}
	\hfill
	\includegraphics[page=5,width=0.45\linewidth]{results1/cr_fm_meta}
	\caption[FM's metagene in Creighton's  data]{Heatmap, box plot and scatter plot showing the association of FM's obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from Creighton's  data, respectively.
	Scales, p-values and $R^2$-value are as described in previous figures.}
	\label{fig:fmmetacr}
\end{figure}

Next, the transformation matrix was applied to the \gls{icgc} cancer data and resulting metagenes were compared with the gene expression and sample \gls{bmi}/\gls{bmi} status.
As evident in \cref{fig:fmmetaicgc}, FM's obesity metagene scores appeared to reflect the overall gene expression of FM's obesity associated genetic signature.
As with all the results in this chapter so far, FM's obesity metagene did not significantly associate with any of the \gls{icgc} cancer data, except in \gls{blca} data set (\cref{fig:fmmetaicgc}).
FM's obesity metagene significantly associated with the overweight group (but not with the obese group), and also had a significant \gls{anova} p-value.
On the contrary to the association of the metagene with the sample \gls{bmi} status, FM's obesity metagene was not associated with sample \gls{bmi}.
These results suggested that the samples that were overweight in the \gls{blca} data set had similar biological properties as the samples that were obese in FM's data set.
However, due to the fact that FM's metagene lacked association with sample \gls{bmi} in \gls{blca} data set and that the metagene did not show any significant association in any other cancer type, it was difficult to determine whether the observed association of FM's obesity metagene with the overweight group of samples was truly reflective of the effect resulted from FM's metagene.
Lastly, the transformation matrix was applied to Print's data (\cref{fig:fmmetacris}).
Again, FM's obesity metagene scores reflected the gene expression of the samples, did not significantly associate with sample \gls{bmi}/\gls{bmi} status.
\\

\begin{figure}[htp!]
	\centering
	\includegraphics[page=3,width=0.8\linewidth]{results1/fm_meta_ICGC}\\
	\vspace{1em}
	\includegraphics[page=4,width=0.45\linewidth]{results1/fm_meta_ICGC}
	\hfill
	\includegraphics[page=5,width=0.45\linewidth]{results1/fm_meta_ICGC}
	\caption[FM's metagene in \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of FM's obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from \acrshort{icgc} \acrshort{blca} data, respectively.
	The results for other cancer types are in \cref{app:a}.
	Scales, p-values and $R^2$-value are as described in previous figures.}
	\label{fig:fmmetaicgc}
\end{figure}

\begin{figure}[htp!]
	\centering
	\includegraphics[page=3,width=0.8\linewidth]{results1/cris_fm_meta}\\
	\vspace{1em}
	\includegraphics[page=4,width=0.45\linewidth]{results1/cris_fm_meta}
	\hfill
	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_fm_meta}
	\caption[FM's metagene in Print's breast cancer data]{Heatmap, box plot and scatter plot showing the association of FM's obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from Print's breast cancer data, respectively.
	Scales, p-values and $R^2$-value are as described in previous figures.}
	\label{fig:fmmetacris}
\end{figure}

\noindent
These results showed that FM's obesity associated metagene was not transferable to other cancer data sets, similar to Creighton's obesity associated metagene.
This meant that both Creighton's and FM's obesity associate metagenes may have been too specific to the orignial data set in which the signatures were identified in.
Furthermore, there was a possibility that these obesity associated metagenes were not related to obesity, but associated with a different clinical variable that may be closely related to \gls{bmi}.

@@

Both the obesity associated metagenes generated from Creighton's data and FM's data were able to capture the overall pattern of gene expression of the genetic signature in the samples, but did not associate with sample \gls{bmi}/\gls{bmi} status.
One possible reason for this result could be that the obesity associated genetic signatures from Creighton's and FM's study may not have been truly associated with sample \gls{bmi} status, but with another clinical variable.
To reject this possibility, obesity associated genetic signatures were identified in Creighton's data after controlling for all the clinical variables in the data set.
FM's data set was not used to get obesity associated genetic signatures, as no \gls{bmi} information was available for the samples in FM's data.

Firstly, to validate whether the identified obesity associated genetic signature was similar to the signature that Creighton's groups had originally found, the identified \glspl{deg} were compared with the set of genes identified by Creighton's group.
\glspl{deg}  were identified between the samples from obese patient and non-obese patient in the \gls{rma}-normalised Creighton's data, as described in \cref{sec:gene_expression_analysis}.
Without adjusting the p-value with FDR, 5278 gene probes and 1781 gene probes were significant at p \textless{} 0.05 and p \textless{} 0.01, respectively.
After afjustment, there were only 9 gene probes significant at p \textless{} 0.05 and no gene was significant at p \textless{} 0.01.
Furthermore, there were only 61 gene probes that were significantly differentially expressed at p \textless{} 0.05 with a log$_2$ \gls{fc} greater than 1.2.
From these observations, the log$_2$ fold change of the gene probes were ignored and the threshold p-value was set to 0.01 (without adjustment) for the identification of gene probes.
This was to include as many gene probes as possible without losing power (?).
Additionally, when there were more than 799 gene probes identified, only the most significant 799 gene probes were taken as the obesity associated genetic signature, as this many genes were originally identified by Creighton's group.
These criteria were applied for the identification of other genetic signatures as well.

The above analysis was repeated with the residual data (\gls{rma}-normalised Creighton's data that had been controlled for other clinical variables; see \cref{sub:residual_data_creation}).
The clinical variables controlled were age, ethnicity, menopause status, tumour grade, hormone (\gls{er}, \gls{pr}, and \gls{her2}) statuses  and \gls{ln} status.
In the residual data, 1104 gene probes were significant with unadjusted p-value (p \textless{} 0.01).
Again, the most significant 799 genes were taken as the obesity associated genetic signature from this data.

In addition to the above two obesity associated genetic signatures, two more sets of genetic signatures were identified by taking the common genes between the above two signatures with Creighton's original obesity associated genetic signatures.
There were 239 common gene probes between Creighton's original genetic signature and the gene probes identified from the unadjusted Creighton's data, and 168 common gene probes between the original signature and gene probes from residual data in Creighton's data.
The genetic signatures identified were summarised as a Venn diagram, as shown in \cref{fig:venn1}.
\\

\begin{figure}[htp!]
	\centering
	\includegraphics[width=0.5\linewidth]{results1/deg_venn1}
	\caption[Veen diagram of the \glspl{deg} identified from Creighton's data (all samples)]{Venn diagram showing the common genes between the signatures obtained from unadjusted and clinical variable-adjusted Creighton's data and Creighton's original obesity associated genetic signature.}
	\label{fig:venn1}
\end{figure}

>>>>

The obesity associated genetic signatures are central to this project in order to clarify the relationship between \gls{bmi} and cancer.
Firstly in this chapter, the previously identified obesity associated signatures from the studies conducted by \citet{Creighton2012} and \citet{Fuentes-Mattei2014} were examined in turn to judge the agreement of these signatures with the sample \gls{bmi}/\gls{bmi} status, as presented in thier results.
Secondly, novel obesity associated genetic signatures were identified in the Creighton \textit{et al.} (CR) data set and was compared with the signatures from the Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} studies.
Lastly, the presence of common genes or pathways that were associated with obesity in multiple types of cancer was explored using the \gls{icgc} cancer data sets.

@@

Obesity metagene was created using the obesity associated genetic signatures from the \citet{Creighton2012} study.
There was no description about the normalisation method used by Creighton \textit{et al.} when they first analysed their data, so the more reliable \gls{rma} method was used to  normalise the CR data set.
The obesity metagene was plotted in a heatmap with the sample gene expression  to check whether the metagene scores were in accordance with the overall gene expression of the samples (\cref{fig:crmetaheat}).

\begin{figure}[htb]
	\centering
	\includegraphics[page=3,width=0.8\linewidth]{results1/creighton_mg_heatmap1}
	\caption[Obesity metagene from \citet{Creighton2012} study and sample gene expression in CR data]{Heatmap showing the obesity metagene from the \citet{Creighton2012} study with the gene expressions of obesity associated genes in CR data.
	Level of expression is represented in the top right histogram, where low and high gene expression were colour-coded with blue and red, respectively.
	Each row of the heatmap represents a gene from the obesity associated genetic signature, and each column of the heatmap represents a sample from CR data.
	The obesity associated metagene scores of the samples are shown in a separate row at top of the heatmap, and the tree diagram of the heirarchical clustering of the genes is shown to the left of the heatmap.
	For clarity, the metagene scores were flipped in order to match the results from the \citet{Creighton2012} study.}
	\label{fig:crmetaheat}
\end{figure}

@@

\begin{figure}[htp!]
	\centering
	\includegraphics[page=4,width=0.45\linewidth]{results1/creighton_mg_heatmap1}
	\hfill
	\includegraphics[page=5,width=0.45\linewidth]{results1/creighton_mg_heatmap1}
	\caption[Obesity metagene from the \citet{Creighton2012} study and sample \gls{bmi}/\gls{bmi} status in CR data]{Box plot and scatter plot showing the association of the obesity metagene from the \citet{Creighton2012} study with sample \gls{bmi} status and \gls{bmi} from CR data, respectively.
	In the box plot, the p-values above the groups represent the statistical significance of the association of the metagene with the overweight or obese group compared with the normal weight group.
	The \gls{anova} p-value shows the statistical significance of the association of the metagene with the sample \gls{bmi} groups.
	In the scatter plot, $R^2$- and p-values describe the adjusted coefficient of determination of the regression line and the statistical significance of the linear model used to draw the regression line, respectively.}
	\label{fig:crmetaboxplot}
\end{figure}

\cref{fig:crmetaboxplot} clearly showed that the obesity metagene from the \citet{Creighton2012} study significantly associated with the sample \gls{bmi} status, as well as sample \gls{bmi} value (include p-value and r-squared).
It should be noted here that the obesity associated metagene significantly associated with the samples that were obese, but not with the samples that were overweight.
This was due to the fact that the obesity associated genetic signatures were originally identified from the comparison of the samples that were obese with the samples that were not obese, and therefore the metagene scores were significant with the obese group, but not with the overweight group.
In addition to this, although the regression line in the scatter plot showed statistically significant association with sample \gls{bmi}, the sample \gls{bmi} values seemed to be randomly dispersed across the metagene scores.
This suggested that perhaps the obesity metagene from the \citet{Creighton2012} study was not associated with sample \gls{bmi} as strongly as expected.
\\

\noindent
Now that the association of the obesity metagene from the \citet{Creighton2012} study was established in CR data set, the obesity associated signature was generated in the \gls{icgc} cancer data.
The direction of the obesity associated metagene was checked in the CR data first, so that high metagene scores reflected high sample \gls{bmi} and low metagene scores reflected low sample \gls{bmi} (\cref{fig:crmetaboxplot}).
The transformation matrix was then created in CR data, as described in \cref{sub:svd}.

All of the \gls{icgc} data were normalised as described in \cref{ssub:rna_seq_data}.
Before the transformation matrix was applied to the log$_{10}$-normalised cancer data, the suitability of standardised data or untouched (non-standardised) data was determined (see \cref{sec:metagenes_created_from_raw_data_vs_standardised_data_icgc}).
From these results, the standardised data was found to be the most suitable to use for matrix transformation.
The transformation matrix was applied to each cancer data set in turn to obtain the Creighton \textit{et al.} obesity metagenes from each of the data set.
Each obesity metagenes were plotted in a heatmap with the corresponding data set in which the metagene was taken from (\cref{fig:crmetaicgc}; \cref{sec:rest_of_the_icgc_cancer_heatmap_resutls}).
These heatmaps confirmed that the obesity associated metagene derived from CR data was able to capture the overall gene expression pattern, where the metagene scores reflected the expression levels of majority of the genes in the signature.
As before, association of the obesity metagene with the sample \gls{bmi} and \gls{bmi} status was checked in their respective cancer data set (\cref{fig:crmetaicgc}; \cref{sec:rest_of_the_icgc_cancer_heatmap_resutls}).
Out of all the cancer types, only \gls{blca} data set showed significant association of obesity metagene with the overweight group (but not with the obese group).
However, neither the \gls{anova} p-value nor the regression line in the scatter plot were statistically significant, suggesting that this apparent association with the overweight group was not due to the obesity metagene (?).

\begin{figure}[htp!]
	\centering
	\includegraphics[page=3,width=0.8\linewidth]{results1/crtcga_std}\\
	\vspace{1em}
	\includegraphics[page=4,width=0.45\linewidth]{results1/crtcga_std}
	\hfill
	\includegraphics[page=5,width=0.45\linewidth]{results1/crtcga_std}
	\caption[Obesity metagene from the \citet{Creighton2012} study in \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of obesity metagene from the \citet{Creighton2012} study with sample gene expression, \gls{bmi} and \gls{bmi} status from \gls{icgc} \gls{blca} data, respectively.
	The results for other cancer types are in \cref{app:a}.
	Scales, p-values and $R^2$-value are as described in previous figures.}
	\label{fig:crmetaicgc}
\end{figure}

There could be several reasons for the apparent lack of association of the metagene with sample \gls{bmi}/\gls{bmi} status.
First, transformation matrix was derived from CR microarray data, but the \gls{icgc} cancer data were \gls{rnaseq} data.
Though the log$_{10}$ normalisation and standardisation of the data was the most appropriate adjustment to be made to the \gls{rnaseq} data, this adjustment was not equivalent to the \gls{rma} normalisation method that was used on the microarray data.
Secondly, none of the \gls{icgc} cancer originated from breast as in CR data.
Since the obesity associated signature was identified in breast cancer data, the signature may be specific to breast cancer and may not be suitable in other cancer types.
\\

\noindent
To check whether the metagene was specific to breast cancer microarray data, the same transformation matrix was applied to the \gls{nzbc} microarray data from \citet{Print2016} study (referred to as \gls{nzbc} data hereafter).
\gls{nzbc} data was normalised with \gls{rma} method and the transformation matrix was applied to the normalised data to obtain the metagene.
The metagene was again compared with the gene expression of the samples with a heatmap and the association of the metagene with the sample \gls{bmi}/\gls{bmi} status was examined with box and scatter plots (\cref{fig:crmetaprint}).

The obesity metagene managed to reflect the overall gene expression of the samples in \gls{nzbc} data.
However, as with \gls{icgc} cancer data, the obesity metagene scores did not significantly associate with the sample \gls{bmi} or \gls{bmi} status.
These results confirmed that the lack of association of the obesity metagene from CR data was not due to the technology in which the data was gathered (mcroarray or \gls{rnaseq}), nor the cancer type in which the transformation matrix was applied to.

\begin{figure}[htp!]
	\centering
	\includegraphics[width=0.8\linewidth,page=3]{results1/cris_cr_trans_meta}\\
	\vspace{1em}
	\includegraphics[width=0.45\linewidth,page=4]{results1/cris_cr_trans_meta}
	\hfill
	\includegraphics[width=0.45\linewidth,page=5]{results1/cris_cr_trans_meta}
	\caption[Obesity metagene from \citet{Creighton2012} study in \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of obesity metagene from the \citet{Creighton2012} study with sample gene expression, \gls{bmi} and \gls{bmi} status from \gls{nzbc} data, respectively.
	Scales, p-values and $R^2$-value are as described in previous figures.}
	\label{fig:crmetaprint}
\end{figure}

All together, all of these results suggest that the obesity metagene identified in the study conducted by \citet{Creighton2012} associated with sample \gls{bmi} and \gls{bmi} status only in CR data.
The lack of association with \gls{bmi} was not due to the type of technology platform in which the data was gathered, as neither the \gls{icgc} \gls{rnaseq} data nor \gls{nzbc} microarray data showed no significant association with the obesity metagene.
Furthermore, the obesity metagene was not dependent on the cancer type since the obesity associated genetic signature did not show significant association in neither the \gls{icgc} cancer data nor in \gls{nzbc} data.

One possible reason why the obesity metagene from the \citet{Creighton2012} study did not show significant association with other data sets could be because the genetic signature was in fact not an obesity specific signature, but a signature that was detected due to another clinical variable.
Another reason for this apparent lack of association could be that the genetic signature was too specific to the data and was not a broad obesity associated genetic signature, but an obesity associated signature specifically for CR data set.

@@

Since the obesity associated genetic signature from the \citet{Creighton2012} study was not associated with sample \gls{bmi} or \gls{bmi} status in other data sets, the obesity associated genetic signature from the \citet{Fuentes-Matter2014} study (FM) was examined to see whether FM obesity metagene associated with sample \gls{bmi} or \gls{bmi} status.
Since FM data set did not have sample \gls{bmi} information, FM obesity metagene was not able to be compared with the sample \gls{bmi} or \gls{bmi} status in the original FM data.
However, the transformation matrix was still applied to other cancer data to see whether FM obesity metagene associated with sample \gls{bmi}/\gls{bmi} status

Firstly, FM data was normalised with the \gls{rma} method and \gls{svd} was applied to the normalised data to get the transformation matrix.
The transformation matrix was used to transform the \gls{rma}-normalised CR data to extract FM obesity metagene scores in CR data.
FM obesity metagene scores were compared with gene expressions and sample \gls{bmi}/\gls{bmi} status in CR data, as shown in \cref{fig:fmmetacr}.
Clearly, like with the obesity metagene identified by Creighton \textit{et al.}, FM obesity metagene was reflective of the overall gene expression of the samples, but did not associate with the sample \gls{bmi} or \gls{bmi} status.

\begin{figure}[htp!]
	\centering
	\includegraphics[page=3,width=0.8\linewidth]{results1/cr_fm_meta}\\
	\vspace{1em}
	\includegraphics[page=4,width=0.45\linewidth]{results1/cr_fm_meta}
	\hfill
	\includegraphics[page=5,width=0.45\linewidth]{results1/cr_fm_meta}
	\caption[FM metagene in CR data]{Heatmap, box plot and scatter plot showing the association of FM obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from CR  data, respectively.
	Scales, p-values and $R^2$-value are as described in previous figures.}
	\label{fig:fmmetacr}
\end{figure}

Next, the transformation matrix was applied to the \gls{icgc} cancer data and resulting metagenes were compared with the gene expression and sample \gls{bmi}/\gls{bmi} status.
As evident in \cref{fig:fmmetaicgc}, FM obesity metagene scores appeared to reflect the overall gene expression of FM obesity associated genetic signature.
As with all the results in this chapter so far, FM obesity metagene did not significantly associate with any of the \gls{icgc} cancer data, except in \gls{blca} data set (\cref{fig:fmmetaicgc}).
FM obesity metagene significantly associated with the overweight group (but not with the obese group), and also had a significant \gls{anova} p-value.
On the contrary to the association of the metagene with the sample \gls{bmi} status, FM obesity metagene was not associated with sample \gls{bmi}.
These results suggested that the samples that were overweight in the \gls{blca} data set had similar biological properties as the samples that were obese in FM data set.
However, due to the fact that FM metagene lacked association with sample \gls{bmi} in \gls{blca} data set and that the metagene did not show any significant association in any other cancer type, it was difficult to determine whether the observed association of FM obesity metagene with the overweight group of samples was truly reflective of the effect resulted from FM metagene.
Lastly, the transformation matrix was applied to \gls{nzbc} data (\cref{fig:fmmetacris}).
Again, FM obesity metagene scores reflected the gene expression of the samples, did not significantly associate with sample \gls{bmi}/\gls{bmi} status.
\\

\begin{figure}[htp!]
	\centering
	\includegraphics[page=3,width=0.8\linewidth]{results1/fm_meta_ICGC}\\
	\vspace{1em}
	\includegraphics[page=4,width=0.45\linewidth]{results1/fm_meta_ICGC}
	\hfill
	\includegraphics[page=5,width=0.45\linewidth]{results1/fm_meta_ICGC}
	\caption[FM metagene in \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of FM obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from \acrshort{icgc} \acrshort{blca} data, respectively.
	The results for other cancer types are in \cref{app:a}.
	Scales, p-values and $R^2$-value are as described in previous figures.}
	\label{fig:fmmetaicgc}
\end{figure}

\begin{figure}[htp!]
	\centering
	\includegraphics[page=3,width=0.8\linewidth]{results1/cris_fm_meta}\\
	\vspace{1em}
	\includegraphics[page=4,width=0.45\linewidth]{results1/cris_fm_meta}
	\hfill
	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_fm_meta}
	\caption[FM metagene in \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of FM obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from \gls{nzbc} data, respectively.
	Scales, p-values and $R^2$-value are as described in previous figures.}
	\label{fig:fmmetacris}
\end{figure}

\noindent
These results showed that FM obesity associated metagene was not transferable to other cancer data sets, similar to the obesity metagene identified by Creighton \textit{et al.}.
This meant that both the obesity metagenes identified by Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} may have been too specific to the orignial data set in which the signatures were identified in.
Furthermore, there was a possibility that these obesity associated metagenes were not related to obesity, but associated with a different clinical variable that may be closely related to \gls{bmi}.

@@

Both the obesity associated metagenes generated from CR data and FM data were able to capture the overall pattern of gene expression of the genetic signature in the samples, but did not associate with sample \gls{bmi}/\gls{bmi} status.
One possible reason for this result could be that the obesity associated genetic signatures from the \citet{Creighton2012} and \citet{Fuente-Matter2014} studies may not have been truly associated with sample \gls{bmi}  status, but with another clinical variable.
To reject this possibility, obesity associated genetic signatures were identified in CR data after controlling for all the clinical variables in the data set.
FM data set was not used to get obesity associated genetic signatures, as no \gls{bmi} information was available for the samples in FM data.

Firstly, to validate whether the identified obesity associated genetic signature was similar to the signature that Creighton \textit{et al.} had originally found, the identified \glspl{deg} were compared with the original genetic signature identified by Creighton \textit{et al.}.
\glspl{deg}  were identified between the samples from obese patient and non-obese patient in the \gls{rma}-normalised CR data, as described in \cref{sec:gene_expression_analysis}.
Without adjusting the p-value with FDR, 5278 gene probes and 1781 gene probes were significant at p \textless{} 0.05 and p \textless{} 0.01, respectively.
After afjustment, there were only 9 gene probes significant at p \textless{} 0.05 and no gene was significant at p \textless{} 0.01.
Furthermore, there were only 61 gene probes that were significantly differentially expressed at p \textless{} 0.05 with a log$_2$ \gls{fc} greater than 1.2.
From these observations, the log$_2$ fold change of the gene probes were ignored and the threshold p-value was set to 0.01 (without adjustment) for the identification of gene probes.
This was to include as many gene probes as possible without losing power (?).
Additionally, when there were more than 799 gene probes identified, only the most significant 799 gene probes were taken as the obesity associated genetic signature, as this many genes were originally identified by Creighton \textit{et al.}.
These criteria were applied for the identification of other genetic signatures as well.

The above analysis was repeated with the residual data (\gls{rma}-normalised CR data that had been controlled for other clinical variables; see \cref{sub:residual_data_creation}).
The clinical variables controlled were age, ethnicity, menopause status, tumour grade, hormone (\gls{er}, \gls{pr}, and \gls{her2}) statuses  and \gls{ln} status.
In the residual data, 1104 gene probes were significant with unadjusted p-value (p \textless{} 0.01).
Again, the most significant 799 genes were taken as the obesity associated genetic signature from this data.

In addition to the above two obesity associated genetic signatures, two more sets of genetic signatures were identified by taking the common genes between the above two signatures with the original obesity associated genetic signatures.
There were 239 common gene probes between the original genetic signature and the gene probes identified from the unadjusted CR data, and 168 common gene probes between the original signature and gene probes from residual data in CR data.
The genetic signatures identified were summarised as a Venn diagram, as shown in \cref{fig:venn1}.
\\

\begin{figure}[htp!]
	\centering
	\includegraphics[width=0.5\linewidth]{results1/deg_venn1}
	\caption[Veen diagram of the \glspl{deg} identified from CR data (all samples)]{Venn diagram showing the common genes between the signatures obtained from unadjusted and clinical variable-adjusted CR data and the original obesity associated genetic signature.}
	\label{fig:venn1}
\end{figure}
